---
title: "R_Basics"
author: "- Vedanti Borate"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Installation and Pre-requistes
To set up and configure R programming using RStudio, follow these steps:

1. **Install R**: First, you need to install R on your system. You can download it from the Comprehensive R Archive Network (CRAN) at https://cran.r-project.org/. Choose the version that matches your operating system (Windows, MacOS, Linux), download it, and follow the installation instructions.

2. **Install RStudio**: RStudio is an integrated development environment (IDE) for R. It provides a user-friendly interface and tools to help you use R more effectively. You can download RStudio from https://www.rstudio.com/products/rstudio/download/. Again, choose the version that matches your operating system, download it, and follow the installation instructions.

3. **Configure RStudio**: After installing RStudio, you can configure it to suit your needs. Here are some basic configurations you might want to make:

   - **Set your working directory**: This is the directory where RStudio will look for files and save your work. You can set it by going to `Tools -> Global Options -> General -> Default working directory`.

   - **Choose your CRAN mirror**: This is the server from which you'll download packages. You can set it by going to `Tools -> Global Options -> Packages -> CRAN mirror`.

   - **Manage your packages**: R packages are collections of functions and data sets developed by the community. You can install, update, and manage your packages by going to `Tools -> Install Packages`.

4. **Start using RStudio**: Now you're ready to start using RStudio. You can write your R code in the script editor, and then run it by clicking the `Run` button or pressing `Ctrl+Enter` (or `Cmd+Enter` on a Mac). The results will appear in the console.

### Commenting in R
R doesn’t support Multi-line and Documentation comments. It only supports single-line comments drafted by a ‘#’ symbol.
Multi-line commenting : select lines > control/command + shift + C

### R Operators 
#### Arithmetic Operators
```{r}
vec1 <- c(0, 2)
vec2 <- c(2, 3)

cat("Addition =", vec1 + vec2)
cat("Subtraction =", vec1 - vec2)
cat("multiplication =", vec1 * vec2)
cat("division =" ,vec1 /vec2)
cat("division remainder=" ,vec1 %% vec2)
cat("power=", vec1 ^ vec2)


```

#### Logical Operators
```{r}
cat("power=", vec1 & vec2)
cat("power=", vec1 | vec2)
cat("power=", !vec1)

#alternatives
cat("and op=", vec1[1] & vec2[1])
cat("and and op=", vec1[1] && vec2[1])
cat("OR", vec1[1] || vec2[1])

```
#### Assignment Operators
```{r}
vec1 <- c(2:5)
c(2:5) ->> vec2
vec3 <<- c(2:5)
vec4 = c(2:5)
c(2:5) -> vec5


cat ("vector 1 :", vec1, "\n")
cat("vector 2 :", vec2, "\n")
cat ("vector 3 :", vec3, "\n")
cat("vector 4 :", vec4, "\n")
cat("vector 5 :", vec5)
```
### Data types
```{r}
# Logical
print(class(TRUE))

# Integer
print(class(3L))

# Numeric
print(class(10.5))

# Complex
print(class(1+2i))

# Character
print(class("12-04-2020"))

#type verification
#is.data_type(object)
print(is.numeric("12-04-2020"))

#type conversion
#as.data_type(object) 
print(as.numeric(TRUE))

```

### Variable methods
```{r}
#finding datatype of variables
print(class("Hi"))

#listing variables in workspace
print(ls())

#remove variables
#rm(variable)

#global local variable
global = 5
display = function(){
  global=2
 
} 
display() 
print(global) #displays outside value
```

### Reading Input
```{r}
#var1 = readline(prompt = "Enter your name : ");
#d = scan(what = double())

```

### Printing Output
```{r}
x <- "GeeksforGeeks"
#single variable
x
#string and variable
print(x)
cat("cat method ",x,9)
paste0("paste method ",x,9)

#displaying message
message(x, " is best") 

#printing decimals
x <- 15 / 7
print(x, digits = 3) 

```

### Control Statments
#### IF-else
```{r}
# if(expression){
#     statements
#     ....
#     ....
# }
var='abc'
if (var == 'abc'){
  print(var)
}else {
  print("not found")
}
```

#### for loop
```{r}
# for(value in vector){
#     statements
#     ....
#     ....
# }

for (val in (1:5)){
  print(val)
}


```

#### while loop
```{r}
# while(expression){
#     statement
#     ....
#     ....
# }

i=1
while(i<10){
  print(i^2)
  i =i+ 1
}
```


#### repeat loop
```{r}

# repeat { 
#    statements
#    ....
#    .... 
#    if(expression) {
#       break
#    }
# }
x = 1

repeat{ 
print(x) 
x = x + 1
if(x > 5){ 
	break
} 
}

```







